---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

This notebook will have some mappings o


```{r, echo=F, warning=F, message=F}
library(ggplot2, warn.conflicts=F, quietly=T)
library(maps, warn.conflicts=F, quietly=T)
library(ggmap, warn.conflicts=F, quietly=T)
library(dplyr, warn.conflicts=F, quietly=T)
library(factoextra, warn.conflicts=F, quietly=T)
library(Hmisc, warn.conflicts=F, quietly=T)
library(gridExtra, warn.conflicts=F, quietly=T)
library(lubridate, warn.conflicts=F, quietly=T)
library(corrplot, warn.conflicts=F, quietly=T)
library(vcd, warn.conflicts=F, quietly=T)
library(gmodels, warn.conflicts=F, quietly=T)

```

The

```{r}
if ("all_month.csv" %in% dir(".") == FALSE) {
  url <- "https://earthquake.usgs.gov/earthquakes/feed/v1.0/summary/all_month.csv"
  download.file(url = url, destfile = "all_month.csv")
}
quakes <- read.csv("all_month.csv", header=TRUE, sep=',', stringsAsFactors = FALSE)
```

How many lines are in our dataset? 

```{r}
  dim(quakes)
```

Increasing date ordering. 

```{r}
quakes = arrange(quakes, -row_number())
```

Lets have a look at the data. 

```{r}
head(quakes)
```

```{r}
tail(quakes)
```

```{r}
str(quakes)
```

## a short description of the variables 

time: time of event
latitude: decimal degrees latitude. Negative values for southern latitudes. Range is [-90.0,90.0]
longitude: decimal degrees longitude. Negative values for western longitudes. [-180.0,180.0]
depth: depth of the event in kilometers
mag: magnitude for the event. Range [-1.0, 10.0]
magType: method or algorithm used to calculate the preferred magnitude for the event
nst: total number of seismic stations used to determine earthquake location
gap: largest azimuthal gap between azimuthally adjacent stations (in degrees)
dmin: horizontal distance from the epicenter to the nearest station (in degrees)
rms: root-mean-square (RMS) travel time residual, in sec, using all weights
net: ID of  data contributor. Identifies the network considered to be the preferred source of information for this event
id: unique identifier for the event. This is the current preferred id for the event, and may change over time
updated: time when the event was most recently updated
place: textual description of named geographic region near to the event. This may be a city name, or a Flinn-Engdahl Region name
type: type of seismic event
horizontalError: uncertainty of reported location of the event in kilometers
depthError: uncertainty of reported depth of the event in kilometers
magError: uncertainty of reported magnitude of the event
magNst: total number of seismic stations used to calculate the magnitude
status: indicates whether the event has been reviewed by a human
locationSource: network that originally authored the reported location of this event
magSource: network that originally authored the reported magnitude for this event

## A sort exploratory analysis

```{r}
(numeric_vars <- names(which(sapply(quakes, class) == "numeric")))

(integer_vars <- names(which(sapply(quakes, class) == "ineger")))

(factor_vars <- names(which(sapply(quakes, class) == "factor")))

(character_vars <- names(which(sapply(quakes, class) == "character")))

```


# Map

Because we have 

```{r}

world <- map_data('world')

title <- paste("Earthquake map from ", paste(quakes$time[1], quakes$time[nrow(quakes)], sep = " to "))

p <- ggplot() + geom_map(data = world, map = world, aes(x = long, y=lat, group=group, map_id=region), fill="white", colour="#7f7f7f", size=0.5)
p <- p + geom_point(data = quakes, aes(x=longitude, y = latitude, colour = mag)) + scale_colour_gradient(low = "#fee8c8",high = "#e34a33") + ggtitle(title)
p

```